---
title: "Upstreamvs.Downstream"
author: "Kathleen Mason"
date: "12/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

##In this Rmarkdown:
#Uploading relevant datasets, Analysis Question 2
#Question 2: Are the DO daily means equivalent upstream and downstream both before and after dam removal?
#Part A of Q2:Are the daily DO means before dam removal equivalent upstream and downstream of the dam?
#Part B of Q2: Are the daily DO means after dam removal equivalent upstream and downstream of the dam?



```{r Setup and Upload}

library(tidyverse)
library(ggridges)
library(patchwork)
library(dplyr)

getwd()

mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

MilburnieDam_AllSitesDO_TimeStampSUBSET<-read_csv("./Processed/MilburnieDam_AllSitesDO_TimeStampSUBSET_processed.csv")
MilburnieDam_AllSitesDO_BeforeSUBSET<-read_csv("./Processed/MilburnieDam_AllSitesDO_AfterSUBSET_processed.csv")
MilburnieDam_AllSitesDO_AfterSUBSET<-read_csv("./Processed/MilburnieDam_AllSitesDO_BeforeSUBSET_processed.csv")


```

#I need to split the BEFORE data set into upstream and downstream to run the two sample t-test on like I did in Question 1
#Then I will also split up the AFTER data into upstream downstream to run the same test again
```{r Setting Up Before Dataset}

#BEFORE DATASET
#Upstream separation
targetup <- c("Upstream1", "Upstream2", "Upstream3")
MilburnieDam_AllSitesDO_BeforeUpstream <- filter(MilburnieDam_AllSitesDO_BeforeSUBSET, Site %in% targetup)
MilburnieDam_AllSitesDO_BeforeUpstream$Location<-"Upstream"
write.csv(MilburnieDam_AllSitesDO_BeforeUpstream, row.names = FALSE, file =("./Processed/MilburnieDam_AllSitesDO_BeforeUpstream_processed.csv"))

#Downstream separation
targetdown <- c("Downstream1", "Downstream2A", "Downstream2B")
MilburnieDam_AllSitesDO_BeforeDownstream <- filter(MilburnieDam_AllSitesDO_BeforeSUBSET, Site %in% targetdown)
MilburnieDam_AllSitesDO_BeforeDownstream$Location<-"Downstream"
write.csv(MilburnieDam_AllSitesDO_BeforeDownstream, row.names = FALSE, file =("./Processed/MilburnieDam_AllSitesDO_BeforeDownstream_processed.csv"))

#Combine the two back together
MilburnieDam_Q2_BeforeUpvsDown<- full_join(MilburnieDam_AllSitesDO_BeforeDownstream, MilburnieDam_AllSitesDO_BeforeUpstream)
write.csv(MilburnieDam_Q2_BeforeUpvsDown, row.names = FALSE, file =("./Processed/MilburnieDam_Q2_BeforeUpvsDown_processed.csv"))


```

#Now we can run a statistical analysis on the Before data to determine if DO means before dam removal are equivalent upstream and downstream of the dam.
```{r Before Removal Statistical Analysis}

shapiro.test(MilburnieDam_Q2_BeforeUpvsDown$AvgDO[MilburnieDam_Q2_BeforeUpvsDown$Location == "Upstream"])
#Results: W = 0.94248, p-value = 3.702e-15

shapiro.test(MilburnieDam_Q2_BeforeUpvsDown$AvgDO[MilburnieDam_Q2_BeforeUpvsDown$Location == "Downstream"])
#Results: W = 0.97528, p-value = 2.667e-11


var.test(MilburnieDam_Q2_BeforeUpvsDown$AvgDO ~ MilburnieDam_Q2_BeforeUpvsDown$Location)
#Results
#F = 1.0381, num df = 908, denom df = 648, p-value = 0.6102
#alternative hypothesis: true ratio of variances is not equal to 1
#95 percent confidence interval:
# 0.8993459 1.1962747
#sample estimates:
#ratio of variances 
#          1.038133 


vartest<-ggplot(MilburnieDam_Q2_BeforeUpvsDown, aes(x = AvgDO, color = Location)) + geom_freqpoly()
vartest

AvgDO.UpvsDown.twosample <- t.test(MilburnieDam_Q2_BeforeUpvsDown$AvgDO ~ MilburnieDam_Q2_BeforeUpvsDown$Location) 
AvgDO.UpvsDown.twosample
#Results
#t = -3.6291, df = 1411.3, p-value = 0.0002945
#alternative hypothesis: true difference in means is not equal to 0
#95 percent confidence interval:
# -0.7957353 -0.2373324
#sample estimates:
#mean in group Downstream   mean in group Upstream 
  #              7.734424                 8.250958 

AvgDO.UpvsDown.twosample$p.value
#Results= 0.0002945363

#Format as GLM
AvgDO.UpvsDown.twosample2 <- lm(MilburnieDam_Q2_BeforeUpvsDown$AvgDO ~ MilburnieDam_Q2_BeforeUpvsDown$Location) 
summary(AvgDO.UpvsDown.twosample2)
plot(AvgDO.UpvsDown.twosample2)



#Normality is not met Pvalue < 0.05

#Wilcoxon Test
AvgDO.UpvsDown.wilcox <- wilcox.test(MilburnieDam_Q2_BeforeUpvsDown$AvgDO ~ MilburnieDam_Q2_BeforeUpvsDown$Location) 

AvgDO.UpvsDown.wilcox
#Results
#W = 257958, p-value = 2.361e-05
#alternative hypothesis: true location shift is not equal to 0

#Interpretation.... significant difference in Avg daily DO during and after the dam removal(p value<0.0001, Wilcoxon test, W=  257958)

#Graph this!
MilburnieDam_Q2_BeforeUpvsDown_Q2boxplot<-ggplot(MilburnieDam_Q2_BeforeUpvsDown) +
geom_boxplot(aes(x = Location , y = AvgDO), color="skyblue4", alpha=0.75)+
  xlab("Location")+
  ylab("Average Dissolved \nOxygen (mgL)")+
  mytheme+
  ggtitle("Before Dam Removal")

MilburnieDam_Q2_BeforeUpvsDown_Q2boxplot

ggsave("./Outputs/MilburnieDam_Q2_BeforeUpvsDown_Q2boxplot.jpg", plot=MilburnieDam_Q2_BeforeUpvsDown_Q2boxplot, width =10, height=6, units="in", dpi=300)

tapply(MilburnieDam_Q2_BeforeUpvsDown$AvgDO, MilburnieDam_Q2_BeforeUpvsDown$Location, summary)


```




#AFTER Now

```{r Setup of After Dataset}
#AFTER DATASET
#Upstream separation
targetup <- c("Upstream1", "Upstream2", "Upstream3")
MilburnieDam_AllSitesDO_AfterUpstream <- filter(MilburnieDam_AllSitesDO_AfterSUBSET, Site %in% targetup)
MilburnieDam_AllSitesDO_AfterUpstream$Location<-"Upstream"
write.csv(MilburnieDam_AllSitesDO_AfterUpstream, row.names = FALSE, file =("./Processed/MilburnieDam_AllSitesDO_AfterUpstream_processed.csv"))

#Downstream separation
targetdown <- c("Downstream1", "Downstream2A", "Downstream2B")
MilburnieDam_AllSitesDO_AfterDownstream <- filter(MilburnieDam_AllSitesDO_AfterSUBSET, Site %in% targetdown)
MilburnieDam_AllSitesDO_AfterDownstream$Location<-"Downstream"
write.csv(MilburnieDam_AllSitesDO_AfterDownstream, row.names = FALSE, file =("./Processed/MilburnieDam_AllSitesDO_AfterDownstream_processed.csv"))

#Combine the two back together
MilburnieDam_Q2_AfterUpvsDown<- full_join(MilburnieDam_AllSitesDO_AfterDownstream, MilburnieDam_AllSitesDO_AfterUpstream)
write.csv(MilburnieDam_Q2_AfterUpvsDown, row.names = FALSE, file =("./Processed/MilburnieDam_Q2_AfterUpvsDown_processed.csv"))

```

#Now we can run a statistical analysis on the After data to determine if DO means before dam removal are equivalent upstream and downstream of the dam.
```{r Statistical Analysis on After Removal}

shapiro.test(MilburnieDam_Q2_AfterUpvsDown$AvgDO[MilburnieDam_Q2_AfterUpvsDown$Location == "Upstream"])
#Results: W = 0.94276, p-value < 2.2e-16

shapiro.test(MilburnieDam_Q2_AfterUpvsDown$AvgDO[MilburnieDam_Q2_AfterUpvsDown$Location == "Downstream"])
#Results: W = 0.96158, p-value < 2.2e-16


var.test(MilburnieDam_Q2_AfterUpvsDown$AvgDO ~ MilburnieDam_Q2_AfterUpvsDown$Location)
#Results
#F = 0.86664, num df = 1342, denom df = 948, p-value = 0.01651
#alternative hypothesis: true ratio of variances is not equal to 1
#95 percent confidence interval:
 #0.7699730 0.9742568
#sample estimates:
#ratio of variances 
 #        0.8666357 


vartest2<-ggplot(MilburnieDam_Q2_AfterUpvsDown, aes(x = AvgDO, color = Location)) + geom_freqpoly()
vartest

AvgDO.UpvsDown2.twosample <- t.test(MilburnieDam_Q2_AfterUpvsDown$AvgDO ~ MilburnieDam_Q2_AfterUpvsDown$Location) 
AvgDO.UpvsDown2.twosample
#Results
#t = -0.84154, df = 1948.4, p-value = 0.4002
#alternative hypothesis: true difference in means is not equal to 0
#95 percent confidence interval:
# -0.2965370  0.1184623
#sample estimates:
#mean in group Downstream   mean in group Upstream 
 #               9.264027                 9.353064 

AvgDO.UpvsDown2.twosample$p.value
#Results= 0.4001508

#Format as GLM
AvgDO.UpvsDown2.twosample2 <- lm(MilburnieDam_Q2_AfterUpvsDown$AvgDO ~ MilburnieDam_Q2_AfterUpvsDown$Location) 
summary(AvgDO.UpvsDown2.twosample2)
plot(AvgDO.UpvsDown2.twosample2)



#***Normality is met Pvalue > 0.05

#Wilcoxon Test
AvgDO.UpvsDown2.wilcox <- wilcox.test(MilburnieDam_Q2_AfterUpvsDown$AvgDO ~ MilburnieDam_Q2_AfterUpvsDown$Location) 

AvgDO.UpvsDown2.wilcox
#Results
#W = 641046, p-value = 0.808
#alternative hypothesis: true location shift is not equal to 0


#Interpretation.... NO significant difference in Avg daily DO during and after the dam removal(p value>0.0001, Wilcoxon test, W= 641046)

#Graph this!
MilburnieDam_Q2_AfterUpvsDown_Q2boxplot<-ggplot(MilburnieDam_Q2_AfterUpvsDown) +
geom_boxplot(aes(x = Location , y = AvgDO), color="skyblue4", alpha=0.75)+
  xlab("Location")+
  ylab("Average Dissolved \nOxygen (mgL)")+
  mytheme+
  ggtitle("After Dam Removal")

MilburnieDam_Q2_AfterUpvsDown_Q2boxplot

ggsave("./Outputs/MilburnieDam_Q2_AfterUpvsDown_Q2boxplot.jpg", plot=MilburnieDam_Q2_AfterUpvsDown_Q2boxplot, width =10, height=6, units="in", dpi=300)


tapply(MilburnieDam_Q2_AfterUpvsDown$AvgDO, MilburnieDam_Q2_AfterUpvsDown$Location, summary)

```

```{r Now using All DO not just Avg}
Upstream1_AllDO<-read_csv("./Processed/MilburnieDam_Upstream1_AllDO_processed.csv")
Upstream2_AllDO<- read_csv("./Processed/MilburnieDam_Upstream2_AllDO_processed.csv")
Impoundment1_AllDO<-read_csv("./Processed/MilburnieDam_Impoundment1_AllDO_processed.csv")
Downstream1_AllDO<-read_csv("./Processed/MilburnieDam_Downstream1_AllDO_processed.csv")
Downstream2A_AllDO<-read_csv("./Processed/MilburnieDam_Downstream2A_AllDO_processed.csv")
Downstream2B_AllDO<-read_csv("./Processed/MilburnieDam_Downstream2B_AllDO_processed.csv")

Upstreams <- full_join(Upstream1_AllDO, Upstream2_AllDO)
UpstreamSites_AllDO<-full_join(Upstreams, Impoundment1_AllDO)
write.csv(UpstreamSites_AllDO, row.names = FALSE, file =("./Processed/MilburnieDam_Q2_UpstreamSites_AllDO_processed.csv"))

Downstreams <- full_join(Downstream1_AllDO, Downstream2A_AllDO)
DownstreamSites_AllDO<-full_join(Downstreams, Downstream2B_AllDO)
write.csv(DownstreamSites_AllDO, row.names = FALSE, file =("./Processed/MilburnieDam_Q2_DownstreamSites_AllDO_processed.csv"))


```

